---
title: "Marquez+2026_PNAS_FiguresCode"
output: html_document
date: "2026-03-27"
---

FIG I. Respirometry Equipment (Fig. 1A), Pcrit Estimation (Fig. 1B) and Brach vs Bivalve comparison (Fig. 1C)

```{r, load the data and create vectors}
exp_data <- read.csv("PaleozoicVModern_PcritSMRData_WarmThermalRangeOnly_ForFigsOnly.csv", header = TRUE)

# remove BLANKs
exp_data <- exp_data %>%
  filter(!(Species %in% c("BLANK")))

kelvin <- 273.15 # kelvin constant
kb <- 8.6173324e-5 # Boltzmann constant
Tref= (1/288.15)*(1/kb)

print(Tref)
```

#################################
Species 1: Terebretalia Transversa
##################################

```{r subset t. transversa data into a separate dataframe, echo=FALSE}
t.transversa_data <- subset(exp_data, Species == "Terebretalia transversa") 
head(t.transversa_data)

t.transversa_data$pcrit <- t.transversa_data$Pcrit_kPa/101.325

ggplot(data = t.transversa_data, aes(x = temp, y = pcrit))+
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, add Arrhenius temp column i.e. 1 / (temp in kelvin * kb)}
t.transversa_data$temp_arrhenius <- 1/((t.transversa_data$temp + kelvin) * kb)
```

```{r, linear model to estimate the temperature sensitivity parameter (E, slope) for Arrhenius function (exp[E/[kb.T]])}
arrhenius_mod <- lm(log(pcrit) ~ temp_arrhenius, data = t.transversa_data)

ggplot(data = t.transversa_data, aes(x = temp_arrhenius, y = log(pcrit)))+
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, get E parameter for Arrhenius function}
arrhenius_E <- arrhenius_mod$coefficients[2]
```

```{r, remove the effect of temperature on pcrit by dividing with the Arrhenius function (exp[E/[kb.T]])}
t.transversa_data$pcrit_temp_norm <- t.transversa_data$pcrit/(exp(arrhenius_E*t.transversa_data$temp_arrhenius))
```

```{r, check that temperature effect is largely removed i.e. slope = close to 0 for pcrit_temp_norm}
ggplot(data = t.transversa_data, aes(x = temp_arrhenius, y = pcrit_temp_norm))+ 
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, estimate mass scaler from power relationship of temperature normalized pcrit vs mass i.e. slope of log - log model}
mass_scaling_lm <- lm(log(pcrit_temp_norm) ~ log(afdw), data = t.transversa_data) 

ggplot(data = t.transversa_data, aes(x = log(afdw), y = log(pcrit_temp_norm)))+ 
  geom_point()+
  geom_smooth(method = 'lm')

# take the slope as the mass scaler
mass_scaler <- -1*as.numeric(mass_scaling_lm$coefficients[2])
```

```{r, mass normalise pcrit data by dividing with mass ^ mass scaler. i.e. B^n in deutsch et al. 2015}
t.transversa_data$pcrit_mass_norm <- t.transversa_data$pcrit*(t.transversa_data$afdw^mass_scaler)

ggplot(data = t.transversa_data, aes(x = temp, y = log(pcrit_mass_norm)))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r, fit a MI model to log(pcrit_mass_norm) v temp_arrhenius. Estimate Ao and Eo}
MI_model <- lm(log(pcrit_mass_norm) ~ temp_arrhenius, data = t.transversa_data) 

Tref_temp_arrhenius <-data.frame(temp_arrhenius=c(Tref))

Ao <- predict(MI_model, Tref_temp_arrhenius)

Eo <- MI_model$coefficients[2]
```

#####################################
Species 2: Glycymeris septentrionalis
#####################################

```{r subset t. transversa data into a separate dataframe, echo=FALSE}
g.septentrionalis_data <- subset(exp_data, Species == "Glycymeris septentrionalis") 
head(g.septentrionalis_data)

g.septentrionalis_data$pcrit <- g.septentrionalis_data$Pcrit_kPa/101.325

ggplot(data = g.septentrionalis_data, aes(x = temp, y = pcrit))+
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, add Arrhenius temp column i.e. 1 / (temp in kelvin * kb)}
g.septentrionalis_data$temp_arrhenius <- 1/((g.septentrionalis_data$temp + kelvin) * kb)
```

```{r, linear model to estimate the temperature sensitivity parameter (E, slope) for Arrhenius function (exp[E/[kb.T]])}
arrhenius_mod_g.septentrionalis <- lm(log(pcrit) ~ temp_arrhenius, data = g.septentrionalis_data)

ggplot(data = g.septentrionalis_data, aes(x = temp_arrhenius, y = log(pcrit)))+
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, get E parameter for Arrhenius function}
arrhenius_E_g.septentrionalis <- arrhenius_mod_g.septentrionalis$coefficients[2]
```

```{r, remove the effect of temperature on pcrit by dividing with the Arrhenius function (exp[E/[kb.T]])}
g.septentrionalis_data$pcrit_temp_norm <- g.septentrionalis_data$pcrit/(exp(arrhenius_E_g.septentrionalis*g.septentrionalis_data$temp_arrhenius))
```

```{r, check that temperature effect is largely removed i.e. slope = close to 0 for pcrit_temp_norm}
ggplot(data = g.septentrionalis_data, aes(x = temp_arrhenius, y = pcrit_temp_norm))+ 
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r, estimate mass scaler from power relationship of temperature normalized pcrit vs mass i.e. slope of log - log model}
mass_scaling_lm_g.septentrionalis <- lm(log(pcrit_temp_norm) ~ log(afdw), data = g.septentrionalis_data) 

ggplot(data = g.septentrionalis_data, aes(x = log(afdw), y = log(pcrit_temp_norm)))+ 
  geom_point()+
  geom_smooth(method = 'lm')

# take the slope as the mass scaler
mass_scaler_g.septentrionalis <- -1*as.numeric(mass_scaling_lm_g.septentrionalis$coefficients[2])

```

```{r, mass normalise pcrit data by dividing with mass ^ mass scaler. i.e. B^n in deutsch et al. 2015}
g.septentrionalis_data$pcrit_mass_norm <- g.septentrionalis_data$pcrit*(g.septentrionalis_data$afdw^mass_scaler_g.septentrionalis)

ggplot(data = g.septentrionalis_data, aes(x = temp_arrhenius, y = log(pcrit_mass_norm)))+
  geom_point()+
  geom_smooth(method = 'lm')

```

```{r, fit a MI model to log(pcrit_mass_norm) v temp_arrhenius. Estimate Ao and Eo}
MI_model_g.septentrionalis <- lm(log(pcrit_mass_norm) ~ temp_arrhenius, data = g.septentrionalis_data) 

Tref_temp_arrhenius <-data.frame(temp_arrhenius=c(Tref))

Ao_g.septentrionalis <- predict(MI_model_g.septentrionalis, Tref_temp_arrhenius)
Eo_g.septentrionalis <- MI_model_g.septentrionalis$coefficients[2]

```

```{r}
# Read images
img_clam <- readPNG("phylopic.clam.616X490.png")
img_brach <- readPNG("phylopic_rhynchonelliformea_192x192.png")
```

```{r, JAMarquez et. al., Brachiopods and Bivalves physiological data comparison}

library(rphylopic)
library(ggplot2)
library(png)
library(grid)
library(ggplot2)
library(png)
library(grid)

shared_theme <- theme_classic() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color = "black"),
    panel.background = element_blank(),
    axis.title = element_text(family = "Helvetica", color = "black", size = 24),
    axis.text  = element_text(family = "Helvetica", color = "black", size = 12),
    legend.position = "none",
    aspect.ratio = 0.6667 
  )

head(t.transversa_data)
head(g.septentrionalis_data)

organismal_comparison <- rbind(t.transversa_data, g.septentrionalis_data)
head(organismal_comparison)
tail(organismal_comparison)

organismal_comparison_Marquez <- ggplot(data = organismal_comparison, aes(x = temp, y = log(pcrit), color=Species))+
  geom_point()+
  scale_color_manual(values = c("Glycymeris septentrionalis"="deepskyblue2", 
                                "Terebretalia transversa"="coral"))+
  geom_smooth(method = 'lm', se = TRUE)+
  shared_theme + 
  labs(
  x = expression(Temperature~degree*C),
  y = expression(Hypoxia~Tolerance~log(P[crit])))

print(organismal_comparison_Marquez)

Plot_1C <- organismal_comparison_Marquez + 
  add_phylopic(img_clam, alpha = 1, x = 8.5, y = -3.45, ysize = .3, color = "black") +
  add_phylopic(img_brach, alpha = 1, x = 15.8, y = -4.4, ysize = .4, color = "black")
  
print(Plot_1C)
```

```{r}
library(ggplot2)
library(cowplot)
library(grid)  # needed if you still want to use rectGrob or draw_grob later

# -----------------------------
# Load plots
# -----------------------------
Plot_1B <- readRDS("Plot_1B.rds")
Plot_1C <- readRDS("Plot_1C.rds")

# -----------------------------
# Standardize font sizes + margins
# -----------------------------
Plot_1B <- Plot_1B +
  theme(
    axis.title   = element_text(size = 11),
    axis.text    = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.text  = element_text(size = 8),
    plot.margin  = margin(4, 6, 4, 6)
  )

Plot_1C <- Plot_1C +
  theme(
    axis.title   = element_text(size = 11),
    axis.text    = element_text(size = 8),
    legend.title = element_text(size = 9),
    legend.text  = element_text(size = 8),
    plot.margin  = margin(4, 6, 4, 6)
  )

# -----------------------------
# Assemble right column (B + C)
# Small vertical gap controlled with rel_heights
# -----------------------------
Right_column <- plot_grid(
  NULL, Plot_1B, Plot_1C, NULL,
  ncol = 1,
  align = "v",
  axis = "lr",
  rel_heights = c(0.15, 1, 1, 0.15)  # keeps spacing between B and C
)

# -----------------------------
# Assemble full figure with A on left, Right_column on right
# Small gap between columns via rel_widths
# -----------------------------
Final_Figure1 <- plot_grid(
  Plot_1A, Right_column,
  ncol = 2,
  rel_widths = c(1.1, 1),  # slightly wider right column to maintain spacing
  align = "h"
)

####### Add A, B, C labels and arrows #############

# Create arrow grobs using geom_curve
arrow_layer <- ggplot() +
  geom_curve(aes(x = -0.03, y = 0.4, xend = 0.07, yend = 0.4), 
             curvature = 0,
             arrow = arrow(length = unit(0.04, "npc"),
                           type = "closed", angle = 20),
             color = "blue", size = 2) +
  geom_curve(aes(x = 0.0, y = 0.92, xend = 0.10, yend = 0.92),
             curvature = 0,
             arrow = arrow(length = unit(0.04, "npc"),
                           type = "closed", angle = 20),
             color = "red", size = 2) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_void()

# Overlay arrows on your figure
Final_Figure1 <- ggdraw(Final_Figure1) +
  draw_plot(arrow_layer) +          # add arrows
  draw_label("A", x = 0.01, y = 0.933,
             hjust = 0, vjust = 1,
             fontface = "bold", size = 17, color = "white") +
  draw_label("B", x = 0.53, y = 0.933,
             hjust = 0, vjust = 1,
             fontface = "bold", size = 17) +
  draw_label("C", x = 0.53, y = 0.449,
             hjust = 0, vjust = 0,
             fontface = "bold", size = 17)

# -----------------------------
# Save as vector PDF (no labels)
# -----------------------------
ggsave(
  "Final_Figure1_labeled.pdf",
  Final_Figure1,
  width = 10,
  height = 9,
  device = pdf  # standard vector PDF
)
```

FIG II. Boxplots of Experimental (Ao and Eo), OBIS bigeographic phys traits (Aeco and Eeco), and Field guide phys traits (low O2, Eeco)

```{r}
library(tidyverse)
library(png)
library(rphylopic)
library(cowplot)
library(car)
```

```{r, read in PNG images downloaded from Phylopic}
# Read images
img_clam <- readPNG("phylopic.clam.616X490.png")
img_brach <- readPNG("phylopic_rhynchonelliformea_192x192.png")
image_echinodermata <- readPNG("phylopic_echinodermata_192x192.png")
image_calliostoma <- readPNG("phylopic_calliostoma_192x192.png")
image_anthozoa <- readPNG("phylopic_anthozoa_192x192.png")
image_crinoidea <- readPNG("phylopic_crinoidea_192x192.png")
```

#####################
Experimental Phys Data
#####################

```{r, load in all experimental data and published literature  physiology data}
boxdata <- read.csv("PaleozoicVModern_FinalMIParameterComparison_WarmThermalRangeOnly.csv", header=T)
```

```{r, make a second dataframe of just my experimental data (excluding stelleroidea)}
boxdata2 <- boxdata %>%
  filter(!(Class %in% c("Asteroidea", "Ophiuroidea")))
```

```{r, USING ALL OF MY EXPERIMENTAL DATA AND PUBLISHED LITERATURE DATA}
##Eo Plot##
Eo_Box1 <- ggplot(boxdata, aes(x= Fauna, y=Eo, fill=Fauna)) +
  geom_boxplot(outlier.size = 4, 
               outlier.shape = 1,
               outlier.stroke = 1.2,
               outlier.colour = "black") +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[o]*")"))

##Ao Boxplot##
Ao_Box1 <- ggplot(boxdata, aes(x= Fauna, y=Ao, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[o]*")"))

Eo_Box1
Ao_Box1

Eo_Box1_phylopic <- Eo_Box1 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 1.2, ysize = 0.192) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 1.2, ysize = 0.256) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.27, y = 1.05, ysize = 0.235) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 1.05, ysize = 0.179) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 1.05, ysize = 0.2131) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 1.05, ysize = 0.240)

Ao_Box1_phylopic <- Ao_Box1 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 200, ysize = 30) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 200, ysize = 40) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 175, ysize = 34.5) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 175, ysize = 28) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 175, ysize = 33.25) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 175, ysize = 35.2)

Eo_Box1_phylopic
Ao_Box1_phylopic

```

```{r}
# Now plot them side by side
Experimental_boxplots <- plot_grid(Ao_Box1, Eo_Box1, ncol = 2)
Experimental_boxplots
```

```{r, BOXPLOTS EXCLUDING STELLAROIDEA}
##Eo Plot##
Eo_Box2 <- ggplot(boxdata2, aes(x= Fauna, y=Eo, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[o]*")"))

##Ao Boxplot##
Ao_Box2 <- ggplot(boxdata2, aes(x= Fauna, y=Ao, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[o]*")"))

Eo_Box2
Ao_Box2

Eo_Box2_phylopic <- Eo_Box2 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 1.2, ysize = 0.192) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 1.2, ysize = 0.256) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.27, y = 1.05, ysize = 0.235) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 1.05, ysize = 0.179) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 1.05, ysize = 0.2131) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 1.05, ysize = 0.240)

Ao_Box2_phylopic <- Ao_Box2 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 200, ysize = 30) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 200, ysize = 40) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 175, ysize = 34.5) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 175, ysize = 28) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 175, ysize = 33.25) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 175, ysize = 35.2)

Eo_Box2_phylopic
Ao_Box2_phylopic

```

```{r}
# Now plot them side by side
Experimental_boxplots_NOSTELLA <- plot_grid(Ao_Box2_phylopic, Eo_Box2_phylopic, ncol = 2)
Experimental_boxplots_NOSTELLA
```

#############
OBIS Boxplots
#############

```{r, create an R dataframe for all OBIS data}
boxdata4 <- read.csv("JAMarquez_OBIStraits_BodySizes_EvoFauna_unique.csv", header=T)
```

```{r, cateogrize faunas}
#Sort out all duplicates for the AphialID
boxdata4<- boxdata4 %>%
  distinct(aphiaID, .keep_all = TRUE)

# Set the order of the levels for the EvoFauna factor
boxdata4$EvoFauna <- factor(boxdata4$EvoFauna, levels = c('M', 'P', 'Other'))

```

```{r, create an R dataframe for OBIS data, excluding stelleroidea}
# Exclude Stellaroidea
boxdata5 <- boxdata4 %>%
  filter(!(class.x %in% c("Asteroidea", "Ophiuroidea")))
```

```{r}
##Eeco Plot##
Eeco_Box4 <- ggplot(boxdata4, aes(x= EvoFauna, y=Eeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))


##Ao Boxplot##
Aeco_Box4 <- ggplot(boxdata4, aes(x= EvoFauna, y= Aeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  coord_cartesian(ylim = c(0, 10)) +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[eco]*")"))

Eeco_Box4
Aeco_Box4

```

```{r}
# Now plot them side by side
OBIS_boxplots <- plot_grid(Aeco_Box4, Eeco_Box4, ncol = 2)

OBIS_boxplots
```

```{r, boxplots of OBIS data excluding asteroidea}
##Eeco Plot##
Eeco_Box5 <- ggplot(boxdata5, aes(x= EvoFauna, y=Eeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))


##Ao Boxplot##
Aeco_Box5 <- ggplot(boxdata5, aes(x= EvoFauna, y= Aeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  coord_cartesian(ylim = c(0, 10)) +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[eco]*")"))

Eeco_Box5
Aeco_Box5

```

```{r}
# Now plot them side by side
OBIS_boxplots2_nostella <- plot_grid(Aeco_Box5, Eeco_Box5, ncol = 2)

OBIS_boxplots2_nostella
```

###########################################
Field Guide (Beneath Pacific Tides) Dataset
###########################################

```{r, create an R dataframe for all Field Guide data}
boxdata6 <- read.csv("Field_Guide_Dataset.csv", header=T)
```

```{r, categorize faunas}
Pal.data <- filter(boxdata6, phylum == "Brachiopoda" | 
                         class == "Anthozoa" |
                         class == "Cephalopoda" |
                         class == "Stenolaemata" |
                         class == "Asteroidea" |
                         class == "Ophiuroidea" |
                         class == "Crinoidea" |
                         class == "Ostracoda") 
Pal.data$evol_fauna <- "P"

Mod.data <- filter(boxdata6, class == "Bivalvia" |
                     class == "Gastropoda" |
                     class == "Malacostraca" |
                     class == "Demospongiae" |
                     class == "Gymnolaemata" |
                     class == "Echinoidea")
Mod.data$evol_fauna <- "M"

fieldguide.boxdata <- rbind(Pal.data, Mod.data)
```

```{r}
fieldguide.boxdata$inverse_low_02 <- 1 / fieldguide.boxdata$low_o2
```

```{r, boxplots for all fieldguide data}
Eeco_fieldguide <- ggplot(fieldguide.boxdata, aes(x= evol_fauna, y=Eo, fill=evol_fauna)) + 
  geom_boxplot(
  outlier.shape = 1,
  outlier.size = 4,
  outlier.stroke = 1.2,
  outlier.colour = "black"
) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

##Low O2 Boxplot##
LowO2_fieldguide  <- ggplot(fieldguide.boxdata, aes(x=evol_fauna, y=low_o2, fill=evol_fauna)) + 
  geom_boxplot(
  outlier.shape = 1,
  outlier.size = 4,
  outlier.stroke = 1.2,
  outlier.colour = "black"
) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  scale_y_reverse() +
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y = expression("Lowest Dissolved O"[2]*" (matm)"))


LowO2_fieldguide
```

```{r}
# Now plot them side by side
fieldguide_boxplots <- plot_grid(LowO2_fieldguide, Eeco_fieldguide, ncol = 2)

fieldguide_boxplots
```

```{r, make an additional field guide dataframe excluding Stelleroidea}
# Assuming your dataframe is called df
fieldguide.boxdata2 <- fieldguide.boxdata %>%
  filter(!(class %in% c("Asteroidea", "Ophiuroidea")))

head(fieldguide.boxdata2)
```

```{r, boxplots for field guide data excluding stelleroidea}
Eeco_fieldguide2 <- ggplot(fieldguide.boxdata2, aes(x= evol_fauna, y=Eo, fill=evol_fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

##Low O2 Boxplot##
LowO2_fieldguide2  <- ggplot(fieldguide.boxdata2, aes(x=evol_fauna, y=low_o2, fill=evol_fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  scale_y_reverse() +
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y = expression("Lowest Dissolved O"[2]*" (matm)"))


LowO2_fieldguide2
```

```{r}
# Now plot them side by side
fieldguide_boxplots2 <- plot_grid(LowO2_fieldguide2, Eeco_fieldguide2, ncol = 2)

fieldguide_boxplots2
```

####################################
Combine boxplots into final figures
####################################

```{r}
make_title <- function(text) {
  ggplot() + 
    theme_void() +
    annotate("text", x = 0.5, y = 0.5, label = text,
             family = "Helvetica", size = 17) +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))
}
```

```{r, create final figure}
library(cowplot)

Experimental_row <- plot_grid(
  make_title("Experimental Data"),
  Experimental_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

OBIS_row <- plot_grid(
  make_title("OBIS Biogeographic Data"),
  OBIS_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Fieldguide_row <- plot_grid(
  make_title("Beneath Pacific Tides Fieldguide Data"),
  fieldguide_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Marquez_fig2 <- plot_grid(
  Experimental_row,
  OBIS_row,
  Fieldguide_row,
  ncol = 1
)

Marquez_fig2
```

```{r, create final figure, EXCLUDING STELLEROIDEA DATA}
Experimental_row2 <- plot_grid(
  make_title("Experimental Data"),
  Experimental_boxplots_NOSTELLA,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

OBIS_row2 <- plot_grid(
  make_title("OBIS Biogeographic Data"),
  OBIS_boxplots2_nostella,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Fieldguide_row2 <- plot_grid(
  make_title("Beneath Pacific Tides Fieldguide Data"),
  fieldguide_boxplots2,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Marquez_fig2.2 <- plot_grid(
  Experimental_row2,
  OBIS_row2,
  Fieldguide_row2,
  ncol = 1
)

Marquez_fig2.2
```

