---
title: "Marquez+2026_PNAS_PhysiologicalComparisons_Statistics"
output: html_document
date: "2025-02-24"
---

This script was used to compare experimentally derived Ao and Eo, OBIS derived bigeographic physiological traits, and field guide derived physiological traits between the Paleozoic and Modern Faunas. 

```{r}
library(tidyverse)
library(png)
library(rphylopic)
library(cowplot)
library(car)
```

```{r, read in PNG images downloaded from Phylopic}
# Read images
img_clam <- readPNG("phylopic.clam.616X490.png")
img_brach <- readPNG("phylopic_rhynchonelliformea_192x192.png")
image_echinodermata <- readPNG("phylopic_echinodermata_192x192.png")
image_calliostoma <- readPNG("phylopic_calliostoma_192x192.png")
image_anthozoa <- readPNG("phylopic_anthozoa_192x192.png")
image_crinoidea <- readPNG("phylopic_crinoidea_192x192.png")
```

```{r, USING ALL OF MY EXPERIMENTAL DATA AND PUBLISHED LITERATURE DATA}
boxdata <- read.csv("PaleozoicVModern_FinalMIParameterComparison_WarmThermalRangeOnly_ForFigsOnly.csv", header=T)
print(boxdata)
```

```{r}
tapply(boxdata$Eo, boxdata$Fauna, median, na.rm = TRUE)
```

```{r}
tapply(boxdata$Ao, boxdata$Fauna, median, na.rm = TRUE)
```

```{r}
boxdata %>% count(Fauna)
```

```{r}
# Function to print boxplot stats for a column within a subset (e.g., P or M)
print_boxplot_stats <- function(data, column_name, fauna_type) {
  values <- data %>% 
    filter(Fauna == fauna_type) %>% 
    pull({{ column_name }})
  
  stats <- fivenum(values)  # min, Q1, median, Q3, max
  
  cat("----", column_name, "(", fauna_type, ")", "----\n")
  cat("Min:            ", stats[1], "\n")
  cat("1st Quartile:   ", stats[2], "\n")
  cat("Median:         ", stats[3], "\n")
  cat("3rd Quartile:   ", stats[4], "\n")
  cat("Max:            ", stats[5], "\n")
}
```

```{r}
# Run stats for all combinations: Ao:P, Ao:M, Eo:P, Eo:M
print_boxplot_stats(boxdata, "Ao", "P")
print_boxplot_stats(boxdata, "Ao", "M")
print_boxplot_stats(boxdata, "Eo", "P")
print_boxplot_stats(boxdata, "Eo", "M")

```

```{r, make a dataframe of experimentally-derived data (including stellaroidea)}
# Assuming your dataframe is called df
boxdata2 <- boxdata %>%
  filter(!(Class %in% c("Asteroidea", "Ophiuroidea")))

print(boxdata2)
```

```{r}
tapply(boxdata2$Eo, boxdata2$Fauna, median, na.rm = TRUE)
```

```{r}
tapply(boxdata2$Ao, boxdata2$Fauna, median, na.rm = TRUE)
```

```{r}
# Run stats for all combinations: Ao:P, Ao:M, Eo:P, Eo:M
print_boxplot_stats(boxdata2, "Ao", "P")
print_boxplot_stats(boxdata2, "Ao", "M")
print_boxplot_stats(boxdata2, "Eo", "P")
print_boxplot_stats(boxdata2, "Eo", "M")
```

```{r, make a dataframe of only experimentally-derived data collected as a part of this current study to compare organisms collected using the same SOPs (exclude physiology data acquired from published literature)}
boxdata3 <- boxdata %>%
  filter(Source == "Experiment")

print(boxdata3)
```


##########################################################################################################
Marquez+ 2026 Figure 2: Boxplots Using Experimentally-derived Data (including Asteroidea and Ophiuroidea)
##########################################################################################################

```{r}
range(boxdata$Ao, na.rm = TRUE)
range(boxdata$Eo, na.rm = TRUE)
```

```{r, USING ALL OF MY EXPERIMENTAL DATA AND PUBLISHED LITERATURE DATA}
##Eo Plot##
Eo_Box1 <- ggplot(boxdata, aes(x= Fauna, y=Eo, fill=Fauna)) +
  geom_boxplot(outlier.size = 4, 
               outlier.shape = 1,
               outlier.stroke = 1.2,
               outlier.colour = "black") +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[o]*")"))

#theme(panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        #axis.line = element_line(colour = "black"),
        #axis.ticks = element_line(color = "black"),
        #panel.background = element_blank(), 
        #axis.title = element_text(family = "Times New Roman", color = "black", size = 18),
        #axis.title.x = element_text(face = "bold"), 
        #axis.title.y = element_text(face = "bold"),
        #axis.text = element_text(family = "Times New Roman", color = "black", size = 12),
        #legend.position = "none")+
  #labs(
  #x = expression(Temp~degree*C),
  #y = expression(Hypoxia~Tolerance~(P[crit] %.% B^epsilon)))


##Ao Boxplot##
Ao_Box1 <- ggplot(boxdata, aes(x= Fauna, y=Ao, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[o]*")"))

Eo_Box1
Ao_Box1

Eo_Box1_phylopic <- Eo_Box1 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 1.2, ysize = 0.192) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 1.2, ysize = 0.256) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.27, y = 1.05, ysize = 0.235) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 1.05, ysize = 0.179) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 1.05, ysize = 0.2131) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 1.05, ysize = 0.240)

Ao_Box1_phylopic <- Ao_Box1 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 200, ysize = 30) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 200, ysize = 40) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 175, ysize = 34.5) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 175, ysize = 28) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 175, ysize = 33.25) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 175, ysize = 35.2)

Eo_Box1_phylopic
Ao_Box1_phylopic

```

```{r}
# Now plot them side by side
Experimental_boxplots <- plot_grid(Ao_Box1, Eo_Box1, ncol = 2)
Experimental_boxplots
```

```{r}
ggsave("Figure2_ExperimentalData.pdf", plot= Experimental_boxplots, width=28, height=20)
```

################################################################################
Statistical tests: ALL Experimentally-derived data (including Stelleroidea)
################################################################################

```{r, test for normality}
shapiro.test(boxdata$Eo[boxdata$Fauna == "P"])
shapiro.test(boxdata$Eo[boxdata$Fauna == "M"])

shapiro.test(boxdata$Ao[boxdata$Fauna == "P"])
shapiro.test(boxdata$Ao[boxdata$Fauna == "M"])
```

```{r, test for homogeneity of variance (robust to non-normality)}
leveneTest(Eo ~ Fauna, data = boxdata)
leveneTest(Ao ~ Fauna, data = boxdata)
```

```{r}
wilcox.test(Eo ~ Fauna, data = boxdata)
wilcox.test(Ao ~ Fauna, data = boxdata)
```

########################################################################################################
Figures and Stats EXCLUDING Asteroidea and Ophiuroidea (including all other experimentally-derived data)
########################################################################################################

```{r, test for normality}
shapiro.test(boxdata2$Eo[boxdata2$Fauna == "P"])
shapiro.test(boxdata2$Eo[boxdata2$Fauna == "M"])

shapiro.test(boxdata2$Ao[boxdata2$Fauna == "P"])
shapiro.test(boxdata2$Ao[boxdata2$Fauna == "M"])
```

```{r, test for homogeneity of variance (assumes normality)}
var.test(Eo ~ Fauna, data = boxdata2)
var.test(Ao ~ Fauna, data = boxdata2)
```

```{r, test for homogeneity of variance (robust to non-normality)}
leveneTest(Eo ~ Fauna, data = boxdata2)
leveneTest(Ao ~ Fauna, data = boxdata2)
```

```{r, non-parametric test, does not assume normality}
wilcox.test(Eo ~ Fauna, data = boxdata2)
wilcox.test(Ao ~ Fauna, data = boxdata2)
```

```{r, BOXPLOTS EXCLUDING STELLEROIDEA (INCLUDING ALL OTHER EXPERIMENTALLY DERIVED DATA) }
##Eo Plot##
Eo_Box2 <- ggplot(boxdata2, aes(x= Fauna, y=Eo, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[o]*")"))

#theme(panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        #axis.line = element_line(colour = "black"),
        #axis.ticks = element_line(color = "black"),
        #panel.background = element_blank(), 
        #axis.title = element_text(family = "Times New Roman", color = "black", size = 18),
        #axis.title.x = element_text(face = "bold"), 
        #axis.title.y = element_text(face = "bold"),
        #axis.text = element_text(family = "Times New Roman", color = "black", size = 12),
        #legend.position = "none")+
  #labs(
  #x = expression(Temp~degree*C),
  #y = expression(Hypoxia~Tolerance~(P[crit] %.% B^epsilon)))

##Ao Boxplot##
Ao_Box2 <- ggplot(boxdata2, aes(x= Fauna, y=Ao, fill=Fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[o]*")"))

Eo_Box2
Ao_Box2

Eo_Box2_phylopic <- Eo_Box2 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 1.2, ysize = 0.192) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 1.2, ysize = 0.256) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.27, y = 1.05, ysize = 0.235) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 1.05, ysize = 0.179) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 1.05, ysize = 0.2131) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 1.05, ysize = 0.240)

Ao_Box2_phylopic <- Ao_Box2 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 200, ysize = 30) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 200, ysize = 40) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 175, ysize = 34.5) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 175, ysize = 28) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 175, ysize = 33.25) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 175, ysize = 35.2)

Eo_Box2_phylopic
Ao_Box2_phylopic

```

```{r}
# Now plot them side by side
Experimental_boxplots_NOSTELLA <- plot_grid(Ao_Box2_phylopic, Eo_Box2_phylopic, ncol = 2)
Experimental_boxplots_NOSTELLA
```
```{r}
ggsave("Boxplots_ExperimentalData_NOSTELLA.pdf", plot= Experimental_boxplots_NOSTELLA, width=28, height=20)
```

#####################################################################################################################################
Figures and Plots Looking at Experimental Data Only Collected During This Study (Excluding Physiology Data From Published Literature)
#####################################################################################################################################

```{r, test for normality}
shapiro.test(boxdata3$Eo[boxdata3$Fauna == "P"])
shapiro.test(boxdata3$Eo[boxdata3$Fauna == "M"])

shapiro.test(boxdata3$Ao[boxdata3$Fauna == "P"])
shapiro.test(boxdata3$Ao[boxdata3$Fauna == "M"])

```

```{r, test for homogeneity of variance}
leveneTest(Eo ~ Fauna, data = boxdata3)
leveneTest(Ao ~ Fauna, data = boxdata3)
```

```{r}
wilcox.test(Eo ~ Fauna, data = boxdata3)
wilcox.test(Ao ~ Fauna, data = boxdata3)
```

```{r}
tapply(boxdata3$Eo, boxdata3$Fauna, mean, na.rm = TRUE)
```

```{r}
tapply(boxdata3$Ao, boxdata3$Fauna, mean, na.rm = TRUE)
```

```{r}
boxdata3 %>% count(Fauna)
```

```{r, BOXPLOTS INCLUDING EXPERIMENTAL DATA ONLY (Excluding Published Literature Data) (Including Stelleroidea) }
##Eo Plot##
Eo_Box3 <- ggplot(boxdata3, aes(x= Fauna, y=Eo, fill=Fauna)) + 
  geom_boxplot(outlier.shape = 1) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[o]*")"))

#theme(panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        #axis.line = element_line(colour = "black"),
        #axis.ticks = element_line(color = "black"),
        #panel.background = element_blank(), 
        #axis.title = element_text(family = "Times New Roman", color = "black", size = 18),
        #axis.title.x = element_text(face = "bold"), 
        #axis.title.y = element_text(face = "bold"),
        #axis.text = element_text(family = "Times New Roman", color = "black", size = 12),
        #legend.position = "none")+
  #labs(
  #x = expression(Temp~degree*C),
  #y = expression(Hypoxia~Tolerance~(P[crit] %.% B^epsilon)))


##Ao Boxplot##
Ao_Box3 <- ggplot(boxdata3, aes(x= Fauna, y=Ao, fill=Fauna)) + 
  geom_boxplot(outlier.shape = 1) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[o]*")"))

Eo_Box3
Ao_Box3

Eo_Box3_phylopic <- Eo_Box3 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 1.2, ysize = .23) + 
  add_phylopic(img_brach, alpha = 1, x = "P", y = 1.2, ysize = .25) + 
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 1.0, ysize = .24) + 
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 1.0, ysize = .19) + 
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 1.1, ysize = .22) + 
  add_phylopic(image_crinoidea, alpha = 1, x = 1.81, y = 1.1, ysize = .24)

Ao_Box3_phylopic <- Ao_Box3 + 
  add_phylopic(img_clam, alpha = 1, x = "M", y = 70, ysize = 11) +
  add_phylopic(img_brach, alpha = 1, x = "P", y = 70, ysize = 13) +
  add_phylopic(image_echinodermata, alpha = 1, x = 1.25, y = 62, ysize = 11.5) +
  add_phylopic(image_calliostoma, alpha = 1, x = 0.78, y = 62, ysize = 9) +
  add_phylopic(image_anthozoa, alpha = 1, x = 2.3, y = 64, ysize = 9.5) +
  add_phylopic(image_crinoidea, alpha = 1, x = 1.8, y = 64, ysize = 11.5)
  
Eo_Box3_phylopic
Ao_Box3_phylopic

```

```{r}
# Now plot them side by side
Boxplots_ExperimentalDataONLY <- plot_grid(Ao_Box3, Eo_Box3, ncol = 2)

print(Boxplots_ExperimentalDataONLY)

```

```{r}
ggsave("BoxplotFig_ExperimentalData_NoLit.pdf", plot= Boxplots_ExperimentalDataONLY, width=28, height=20)
```

##################################################################################################
Figures and Statistics Comparing Paleozoic and Modern Faunas using OBIS-derived physiological data
##################################################################################################

```{r}
boxdata4 <- read.csv("JAMarquez_OBIStraits_BodySizes_EvoFauna_unique.csv", header=T)
```

```{r, cateogrize faunas}
#Sort out all duplicates for the AphialID
boxdata4<- boxdata4 %>%
  distinct(aphiaID, .keep_all = TRUE)

# Set the order of the levels for the EvoFauna factor
boxdata4$EvoFauna <- factor(boxdata4$EvoFauna, levels = c('M', 'P', 'Other'))

head(boxdata4)
summary(boxdata4)
```

```{r, remove "Other" fauna}
boxdata4 <- subset(boxdata4, EvoFauna != "Other")

tail(boxdata4)
```

```{r}
boxdata4 %>% count(EvoFauna)
```

```{r}
# Function to print boxplot stats for a column within a subset (e.g., P or M)
print_boxplot_stats <- function(data, column_name, fauna_type) {
  values <- data %>% 
    filter(EvoFauna == fauna_type) %>% 
    pull({{ column_name }})
  
  stats <- fivenum(values)  # Q1, median, Q3
  
  cat("----", column_name, "(", fauna_type, ")", "----\n")
  cat("1st Quartile:   ", stats[2], "\n")
  cat("Median:         ", stats[3], "\n")
  cat("3rd Quartile:   ", stats[4], "\n")
}
```

```{r}
# Run stats for all combinations: Aeco:P, Aeco:M, Eeco:P, Eeco:M
print_boxplot_stats(boxdata4, "Aeco", "P")
print_boxplot_stats(boxdata4, "Aeco", "M")
print_boxplot_stats(boxdata4, "Eeco", "P")
print_boxplot_stats(boxdata4, "Eeco", "M")
```

############################################################
Statistics for OBIS data comparisons INCLUDING STELLEROIDEA
############################################################

```{r, test for normality. Sample size is too large so can't use Shapiro.test()}
shapiro.test(boxdata4$Eeco[boxdata4$EvoFauna == "P"])
shapiro.test(boxdata4$Eeco[boxdata4$EvoFauna == "M"])

shapiro.test(boxdata4$Aeco[boxdata4$EvoFauna == "P"])
shapiro.test(boxdata4$Aeco[boxdata4$EvoFauna == "M"])
```

```{r}
# Perform the test and store the result
shapiro_result_EecoP <- shapiro.test(boxdata4$Eeco[boxdata4$EvoFauna == "P"])

# Print the p-value with higher precision
options(digits = 20)
shapiro_result_EecoP$p.value
```

```{r}
# Perform the test and store the result
shapiro_result_AecoP <- shapiro.test(boxdata4$Aeco[boxdata4$EvoFauna == "P"])

# Print the p-value with higher precision
options(digits = 20)
shapiro_result_AecoP$p.value
```

```{r}
# Histogram
hist(boxdata4$Eeco[boxdata4$EvoFauna == "M"], breaks = 50)

# Q-Q Plot
qqnorm(boxdata4$Eeco[boxdata4$EvoFauna == "M"])
qqline(boxdata4$Eeco[boxdata4$EvoFauna == "M"], col = "red")
```

```{r}
set.seed(123)  # for reproducibility
sample_data <- sample(boxdata4$Eeco[boxdata4$EvoFauna == "M"], 5000)
shapiro_result_EecoM <- shapiro.test(sample_data)

print(shapiro_result_EecoM)
```

```{r}
# Print the p-value with higher precision
options(digits = 20)
shapiro_result_EecoM$p.value
```

```{r}
set.seed(1234)  # for reproducibility
sample_data2 <- sample(boxdata4$Aeco[boxdata4$EvoFauna == "M"], 5000)
shapiro_result_AecoM <- shapiro.test(sample_data2)

print(shapiro_result_AecoM)
```

```{r}
# Print the p-value with higher precision
options(digits = 20)
shapiro_result_AecoM$p.value
```

```{r, test for homogeneity of variance}
leveneTest(Eeco ~ EvoFauna, data = boxdata4)
leveneTest(Aeco ~ EvoFauna, data = boxdata4)
```

```{r}
wilcox_result_Eeco <- wilcox.test(Eeco ~ EvoFauna, data = boxdata4)
wilcox.test(Aeco ~ EvoFauna, data = boxdata4)
```

```{r}
# Print the p-value with higher precision
options(digits = 20)
wilcox_result_Eeco$p.value
```

```{r}
range(boxdata4$Aeco, na.rm = TRUE)
range(boxdata4$Eeco, na.rm = TRUE)
```

```{r}
summary(boxdata4)
is.finite(boxdata4$Aeco)
is.finite(boxdata4$Eeco)
```

```{r}
##Eeco Plot##
Eeco_Box4 <- ggplot(boxdata4, aes(x= EvoFauna, y=Eeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

##Ao Boxplot##
Aeco_Box4 <- ggplot(boxdata4, aes(x= EvoFauna, y= Aeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  coord_cartesian(ylim = c(0, 10)) +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[eco]*")"))

Eeco_Box4
Aeco_Box4

```

```{r}
# Now plot them side by side
OBIS_boxplots <- plot_grid(Aeco_Box4, Eeco_Box4, ncol = 2)

OBIS_boxplots
```

#############################################################
Statistics AND Boxplots for OBIS data EXCLUDING STELLEROIDEA
#############################################################

```{r}
# Exclude Stelleroidea
boxdata5 <- boxdata4 %>%
  filter(!(class.x %in% c("Asteroidea", "Ophiuroidea")))

print(boxdata5)
```

```{r}
# Run stats for all combinations: Aeco:P, Aeco:M, Eeco:P, Eeco:M
print_boxplot_stats(boxdata5, "Aeco", "P")
print_boxplot_stats(boxdata5, "Aeco", "M")
print_boxplot_stats(boxdata5, "Eeco", "P")
print_boxplot_stats(boxdata5, "Eeco", "M")
```

```{r}
shapiro_result_Eeco_P_NoStella <- shapiro.test(boxdata5$Eeco[boxdata5$EvoFauna == "P"])
# Print the p-value with higher precision
options(digits = 20)
shapiro_result_Eeco_P_NoStella$p.value
```

```{r}
shapiro_result_Aeco_P_NoStella <- shapiro.test(boxdata5$Aeco[boxdata5$EvoFauna == "P"])
# Print the p-value with higher precision
options(digits = 20)
shapiro_result_Aeco_P_NoStella$p.value
```

```{r}
set.seed(12345)  # for reproducibility
sample_data3 <- sample(boxdata5$Eeco[boxdata5$EvoFauna == "M"], 5000)
shapiro.test(sample_data3)
```

```{r}
shapiro_result_Eeco_M_NoStella <- shapiro.test(sample_data3)
# Print the p-value with higher precision
options(digits = 20)
shapiro_result_Eeco_M_NoStella$p.value
```

```{r, test for homogeneity of variance}
leveneTest(Eeco ~ EvoFauna, data = boxdata5)
leveneTest(Aeco ~ EvoFauna, data = boxdata5)
```

```{r}
wilcox_result_Eeco_nostella <- wilcox.test(Eeco ~ EvoFauna, data = boxdata5)
wilcox.test(Aeco ~ EvoFauna, data = boxdata5)
```

```{r}
# Print the p-value with higher precision
options(digits = 20)
wilcox_result_Eeco_nostella$p.value
```


```{r}
##Eeco Plot##
Eeco_Box5 <- ggplot(boxdata5, aes(x= EvoFauna, y=Eeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

##Ao Boxplot##
Aeco_Box5 <- ggplot(boxdata5, aes(x= EvoFauna, y= Aeco, fill=EvoFauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  coord_cartesian(ylim = c(0, 10)) +
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Hypoxia Tolerance (A"[eco]*")"))

Eeco_Box5
Aeco_Box5

```

```{r}
# Now plot them side by side
OBIS_boxplots2_nostella <- plot_grid(Aeco_Box5, Eeco_Box5, ncol = 2)

OBIS_boxplots2_nostella
```


##################################################################################
Statistical tests & Boxplots for Field Guide-derived Data (Beneath Pacific Tides)
##################################################################################

```{r}
boxdata6 <- read.csv("Field_Guide_Data.csv", header=T)

head(boxdata6)
```

```{r}
Pal.data <- filter(boxdata6, phylum == "Brachiopoda" | 
                         class == "Anthozoa" |
                         class == "Cephalopoda" |
                         class == "Stenolaemata" |
                         class == "Asteroidea" |
                         class == "Ophiuroidea" |
                         class == "Crinoidea" |
                         class == "Ostracoda") 
Pal.data$evol_fauna <- "P"

Mod.data <- filter(boxdata6, class == "Bivalvia" |
                     class == "Gastropoda" |
                     class == "Malacostraca" |
                     class == "Demospongiae" |
                     class == "Gymnolaemata" |
                     class == "Echinoidea")
Mod.data$evol_fauna <- "M"

fieldguide.boxdata <- rbind(Pal.data, Mod.data)
```

```{r}
fieldguide.boxdata %>% count(evol_fauna)
```

```{r}
# Function to print boxplot stats for a column within a subset (e.g., P or M)
print_boxplot_stats <- function(data, column_name, fauna_type) {
  values <- data %>% 
    filter(evol_fauna == fauna_type) %>% 
    pull({{ column_name }})
  
  stats <- fivenum(values)  # min, Q1, median, Q3, max
  
  cat("----", column_name, "(", fauna_type, ")", "----\n")
  cat("Min:            ", stats[1], "\n")
  cat("1st Quartile:   ", stats[2], "\n")
  cat("Median:         ", stats[3], "\n")
  cat("3rd Quartile:   ", stats[4], "\n")
  cat("Max:            ", stats[5], "\n")
}
```

```{r}
# Run stats for all combinations: Ao:P, Ao:M, Eo:P, Eo:M
print_boxplot_stats(fieldguide.boxdata, "low_o2", "P")
print_boxplot_stats(fieldguide.boxdata, "low_o2", "M")
print_boxplot_stats(fieldguide.boxdata, "Eo", "P")
print_boxplot_stats(fieldguide.boxdata, "Eo", "M")

```

```{r}
fieldguide.boxdata$inverse_low_02 <- 1 / fieldguide.boxdata$low_o2
```

```{r}
Eeco_fieldguide <- ggplot(fieldguide.boxdata, aes(x= evol_fauna, y=Eo, fill=evol_fauna)) + 
  geom_boxplot(
  outlier.shape = 1,
  outlier.size = 4,
  outlier.stroke = 1.2,
  outlier.colour = "black"
) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

##Low O2 Boxplot##
LowO2_fieldguide  <- ggplot(fieldguide.boxdata, aes(x=evol_fauna, y=low_o2, fill=evol_fauna)) + 
  geom_boxplot(
  outlier.shape = 1,
  outlier.size = 4,
  outlier.stroke = 1.2,
  outlier.colour = "black"
) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  scale_y_reverse() +
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y = expression("Lowest Dissolved O"[2]*" (matm)"))

#ggsave("Box & Whisker Plots on Lowest Dissolved O2 (atm) for Evolutionary Fauna.pdf", width=16, height=16, units="in")

LowO2_fieldguide
```

```{r, test for normality}
shapiro.test(fieldguide.boxdata$Eo[fieldguide.boxdata$evol_fauna == "P"])
shapiro.test(fieldguide.boxdata$Eo[fieldguide.boxdata$evol_fauna == "M"])

shapiro.test(fieldguide.boxdata$low_o2[fieldguide.boxdata$evol_fauna == "P"])
shapiro.test(fieldguide.boxdata$low_o2[fieldguide.boxdata$evol_fauna == "M"])

```

```{r, test for homogeneity of variance}
leveneTest(Eo ~ evol_fauna, data = fieldguide.boxdata)
leveneTest(low_o2 ~ evol_fauna, data = fieldguide.boxdata)
```

```{r}
wilcox.test(Eo ~ evol_fauna, data = fieldguide.boxdata)
wilcox.test(low_o2 ~ evol_fauna, data = fieldguide.boxdata)
```

```{r}
# Now plot them side by side
fieldguide_boxplots <- plot_grid(LowO2_fieldguide, Eeco_fieldguide, ncol = 2)

fieldguide_boxplots
```

#################################################################
EXCLUDE STELLEROIDEA FROM BPT FIELD GUIDE DATASET
#################################################################

```{r, make an additional dataframe excluding Stellaroidea}
# Assuming your dataframe is called df
fieldguide.boxdata2 <- fieldguide.boxdata %>%
  filter(!(class %in% c("Asteroidea", "Ophiuroidea")))

head(fieldguide.boxdata2)
```

```{r}
fieldguide.boxdata2 %>% count(evol_fauna)
```

```{r}
# Run stats for all combinations: Ao:P, Ao:M, Eo:P, Eo:M
print_boxplot_stats(fieldguide.boxdata2, "low_o2", "P")
print_boxplot_stats(fieldguide.boxdata2, "low_o2", "M")
print_boxplot_stats(fieldguide.boxdata2, "Eo", "P")
print_boxplot_stats(fieldguide.boxdata2, "Eo", "M")

```

```{r, test for normality}
shapiro.test(fieldguide.boxdata2$Eo[boxdata2$Fauna == "P"])
shapiro.test(fieldguide.boxdata2$Eo[boxdata2$Fauna == "M"])

shapiro.test(fieldguide.boxdata2$low_o2[boxdata2$Fauna == "P"])
shapiro.test(fieldguide.boxdata2$low_o2[boxdata2$Fauna == "M"])

```

```{r, test for homogeneity of variance}
leveneTest(Eo ~ evol_fauna, data = fieldguide.boxdata2)
leveneTest(low_o2 ~ evol_fauna, data = fieldguide.boxdata2)
```

```{r}
wilcox.test(Eo ~ evol_fauna, data = fieldguide.boxdata2)
wilcox.test(low_o2 ~ evol_fauna, data = fieldguide.boxdata2)
```

```{r}
Eeco_fieldguide2 <- ggplot(fieldguide.boxdata2, aes(x= evol_fauna, y=Eo, fill=evol_fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) +  
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y=expression("Temperature Sensitivity (E"[eco]*")"))

#theme(panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        #axis.line = element_line(colour = "black"),
        #axis.ticks = element_line(color = "black"),
        #panel.background = element_blank(), 
        #axis.title = element_text(family = "Times New Roman", color = "black", size = 18),
        #axis.title.x = element_text(face = "bold"), 
        #axis.title.y = element_text(face = "bold"),
        #axis.text = element_text(family = "Times New Roman", color = "black", size = 12),
        #legend.position = "none")+
  #labs(
  #x = expression(Temp~degree*C),
  #y = expression(Hypoxia~Tolerance~(P[crit] %.% B^epsilon)))


##Low O2 Boxplot##
LowO2_fieldguide2  <- ggplot(fieldguide.boxdata2, aes(x=evol_fauna, y=low_o2, fill=evol_fauna)) + 
  geom_boxplot(
    outlier.shape = 1,
    outlier.size = 4,
    outlier.stroke = 1.2,
    outlier.colour = "black"
  ) + 
  scale_fill_manual(values=c("M" = "deepskyblue2",
                             "P" = "darkorange2")) +
  scale_color_manual(values=c("darkgrey",
                              "darkgrey"))+
  scale_y_reverse() +
  theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(color = "black"),
        axis.title = element_text(family = "Helvetica", color = "black", size = 42),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(family = "Helvetica", color = "black", size = 35),
        legend.position = "none")+
  labs(y = expression("Lowest Dissolved O"[2]*" (matm)"))

#ggsave("Box & Whisker Plots on Lowest Dissolved O2 (atm) for Evolutionary Fauna.pdf", width=16, height=16, units="in")

LowO2_fieldguide2
```

```{r}
# Now plot them side by side
fieldguide_boxplots2 <- plot_grid(LowO2_fieldguide2, Eeco_fieldguide2, ncol = 2)

fieldguide_boxplots2
```


#########################################################################
Combine All Boxplots into Marquez+ 2026 Figure 2 (Including Stelleroidea)
#########################################################################

```{r}
library(cowplot)

# Create horizontal pairs (2 columns per row)
Experimental_boxplots <- plot_grid(
  Ao_Box1_phylopic, NULL, Eo_Box1_phylopic,
  ncol = 3,
  rel_widths = c(1, 0.05, 1),
  align = "v"
)

OBIS_boxplots <- plot_grid(
  Aeco_Box4, NULL, Eeco_Box4,
  ncol = 3,
  rel_widths = c(1, 0.05, 1),
  align = "v"
)

fieldguide_boxplots <- plot_grid(
  LowO2_fieldguide, NULL, Eeco_fieldguide,
  ncol = 3,
  rel_widths = c(1, 0.05, 1),
  align = "v"
)

# Stack the rows vertically (3 rows total)
Marquez_25_fig2 <- plot_grid(
  Experimental_boxplots,
  OBIS_boxplots,
  fieldguide_boxplots,
  ncol = 1,
  align = "hv",
  rel_heights = c(1, 1, 1)  # Optional: adjust row heights if needed
)

```

```{r}
make_title <- function(text) {
  ggplot() + 
    theme_void() +
    annotate("text", x = 0.5, y = 0.5, label = text,
             family = "Helvetica", size = 17) +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))
}
```

```{r}
print(image_echinodermata)
print(image_calliostoma)
print(image_anthozoa)
print(image_crinoidea)
```

```{r, add labels to boxplots of final figure}
library(cowplot)

Experimental_row <- plot_grid(
  make_title("Experimental Data"),
  Experimental_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

OBIS_row <- plot_grid(
  make_title("OBIS Biogeographic Data"),
  OBIS_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Fieldguide_row <- plot_grid(
  make_title("Beneath Pacific Tides Fieldguide Data"),
  fieldguide_boxplots,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Marquez_26_fig2 <- plot_grid(
  Experimental_row,
  OBIS_row,
  Fieldguide_row,
  ncol = 1
)

ggsave("Marquez_26_fig2.pdf", plot = Marquez_26_fig2, width=20, height=28)

```

###############################################################
Combine All Boxplots into a figure (Excluding Stellaroidea)
###############################################################

```{r, add labels to boxplots of final figure}
Experimental_row2 <- plot_grid(
  make_title("Experimental Data (Excluding Stellaroidea)"),
  Experimental_boxplots_NOSTELLA,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

OBIS_row2 <- plot_grid(
  make_title("OBIS Biogeographic Data (Excluding Stellaroidea)"),
  OBIS_boxplots2_nostella,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Fieldguide_row2 <- plot_grid(
  make_title("Beneath Pacific Tides Fieldguide Data (Excluding Stellaroidea)"),
  fieldguide_boxplots2,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Marquez_26_figS1_NOSTELLA <- plot_grid(
  Experimental_row2,
  OBIS_row2,
  Fieldguide_row2,
  ncol = 1
)

Marquez_25_figS1_NOSTELLA

ggsave("Marquez_26_figS1_NOSTELLA.pdf", plot= Marquez_26_figS1_NOSTELLA, width=20, height=28)
```

```{r}
library(cowplot)

# Stack the rows vertically (3 rows total)
Marquez_26_fig2 <- plot_grid(
  Experimental_boxplots_NOSTELLA,
  OBIS_boxplots2_nostella,
  fieldguide_boxplots2,
  ncol = 1,
  align = "hv",
  rel_heights = c(1, 1, 1)  # Optional: adjust row heights if needed
)

```

```{r}
make_title <- function(text) {
  ggplot() + 
    theme_void() +
    annotate("text", x = 0.5, y = 0.5, label = text,
             family = "Helvetica", fontface = "bold", size = 25) +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))
}
```

```{r, add labels to boxplots of final figure}
Experimental_row2 <- plot_grid(
  make_title("Experimental Data"),
  Experimental_boxplots_NOSTELLA,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

OBIS_row2 <- plot_grid(
  make_title("OBIS Biogeographic Data"),
  OBIS_boxplots2_nostella,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Fieldguide_row2 <- plot_grid(
  make_title("Beneath Pacific Tides Fieldguide Data"),
  fieldguide_boxplots2,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

Marquez_26_fig2.2 <- plot_grid(
  Experimental_row2,
  OBIS_row2,
  Fieldguide_row2,
  ncol = 1
)

ggsave("Marquez_26_fig2_NOSTELLA.png", plot= Marquez_26_fig2.2, width=20, height=28, dpi= 300)
```
